Policy Implications of Coupled Economic and Ecological Dynamics in Aquatic Systems

### 2022-11-17 03_marlin_paper.qmd
getting there. coral and pelagic case studies up and in paper

---
title: "Efficient Simulation of Economic and Ecological Dynamics in Aquatic Systems"
author:
    - Daniel Ovando
    - Darcy Bradley
    - Echelle Burns
    - Lennon Thomas
    - James Thorson
date: "`r Sys.Date()`"
format:
  pdf:
    toc: true
    fig-pos: 'H'
  html: default
  docx: default
execute:
  echo: false
  message: false
  warning: false
bibliography: references.bib
params:
  run_name: ["v0.1"]
abstract: "Very rough cut. Communities around the world are increasingly looking to spatial strategies for managing marine ecosystems. However, the design of spatial management strategies faces numerous challenges. Marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond to changes in marine ecosystems in highly complex ways. We present a framework for efficient simulation of the bio-economic dynamics of marine systems, designed to support both management and research. We demonstrate the model's capabilities by focusing on three case studies on the conservation and food security impacts of marine protected areas (MPAs)."
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
---

@@

theme_set(marlin::theme_marlin())

@@

rdatas <- results_files[str_detect(results_files,'.Rdata')]

walk(rdatas, ~load(file.path(results_path,.x), envir = .GlobalEnv)) # load results into global environment

@@

`marlin` follows a "models of intermediate complexity" (MICE) philosophy
[@plagányi2014] targeted specifically towards capturing heterogeneous
habitats, movement dynamics, and technical interactions across species.
In its general form, `marlin` simulates the behavior of populations of
spatially explicit biologically independent age-structured animal
populations affected by one or more fishing fleets in time steps
specified by the user (generally annual or finer). While `marlin` does
not simulate species interactions, fishing fleets in `marlin` are
capable of targeting and affecting multiple species simultaneously. This
allows the model to simulate processes such as fisheries bycatch or
effort displacement in a way that accounts for the ways that changes in
fleet behavior might affect multiple species in the system. The ability
to track the impacts of fishing fleets across multiple species
simultaneously is particularly important as very few fishing fleets are
truly single species.

@@

A fishing fleet is defined in `marlin` by a total pool of effort that
can have different catchability, prices, and selectivity for individual
species within a simulation (each of which we denote as a *metier*. For
example, both a longline and purse seine fleet may capture bigeye
(*Thunnus obesus*) and skipjack (*Katswamus pelamus*) tunas, but the
longline fleet is much more likely to catch bigeye than skipjack, and
*vice versa*. Each fishing fleet generates catches, revenues, and
profits from individual species that it targets, and makes decisions
based on its total objective function across all species.

Revenues *R* for fleet *f* in time step *t* are given by

@@

where *p* is patch, *s* is species, *a* is age, $\alpha$ is selectivity
at age, *q* is catchability , *u* is total fishing mortality, $\Pi$ is
price per unit weight, and *Y* is total catches.

@@

Selectivity $\alpha$ is modeled through a logistic form per

@@

where $l$ is length at age, $l^{sel}$ is the length at 50% selectivity,
and $\delta$ is the difference between the length at 50% selectivity and
the length at 95% selectivity.

or through a dome shaped form of

XX add dome shaped equation XX

*z* is total mortality given by

@@

Where *m* is natural mortality.

This means that each species experiences a total mortality at age in a
time step in a patch, individual fractions of which are portioned off as
catches and subsequently revenues for each fleet. Given revenues *R*, we
then calculate profits $\phi$ per

@@

where costs *C* are calculated

@@

where $$\gamma_f$$ the scales the cost per unit effort for fleet *f*,
$\beta$ allows for the cost of effort to scale non-linearly, and $\eta$
is an additional cost per unit effort of fishing in each patch *p*,
allowing for the model to account for travel costs for different
patches. Travel costs are calculated based on the euclidean distance of
each patch to the nearest port of a given fleet, users can specify any
number from zero to the number of patches of port locations. When no
ports are specified, travel costs are zero.

@@

The more complex option is "open access". Under open access, the total
amount of effort in a given time step for fleet *f* is a function of the
profitability of that fleet in the proceeding time steps in which
fishing was open.

@@

There $\theta$ controls the responsiveness of fleet *f* to the log of
the ratio of total revenues *R* to total costs *C*, and approximates the
percentage amount effort changes in response to a one unit changes in
the log revenue to cost ratio.

Three general exceptions exist for the effort dynamics. When MPAs are
implemented, effort that used to operate inside the MPAs can either
leave the fishery, or be redistributed outside the the MPA (the default
behavior). The model can also accept total catch quotas for each species
in each time step. When quotas are activated, if the total catch across
all fleets for a given species under the model's effort dynamics would
exceed the allowable quota for that species, the total effort for each
fleet is decrease proportional to its contribution to the total catch
until the quota is satisfied. `marlin` also allows users to specify
closed fishing seasons for one or more species in the system. During
closed seasons, no effort is allocated to patches that would result in
capture of the covered species, potentially resulting in zero effort if
the species cannot be avoided. Under these conditions, the amount of
effort in the next open season is a function of the effort and profits
in the last open fishing season.

@@

The underlying population model used is an age structured single-species
model in the manner of @ovando2021a.

For the population model, numbers *n* at time *t* for age *a* are given
by

@@

where *BH* is the Beverton-Holt recruitment function, *ssb* is
spawning-stock-biomass.

where $l_a$ is the mean length at at age, $l_sel$ is the length at which
on average 50% of individuals are selected by the fishery, and
$\delta_{sel}$ are the additional units of length at which on average
95% of fish are selected by the fishery.

*ssb* is calculated by converting age to mean length, calculating weight
at age, maturity at age, and then calculating spawning stock biomass as
the sum of spawning potential at age in a given time step.

$$
  l_{a,s} = l_{\infty,s}\left(1 - e^{-k_s(a - a0_{s})}\right)
$$ {#eq-len}

Biomass *b* at age is then given by

@@

and maturity *mat* is calculated as

$$
  mat_{a,s} = \frac{1}{(1 + e^{-log(19)\times\frac{l_{a,s} - lmat_s}{{\delta}mat_s}}}
$$ {#eq-mat)}

where $l_{mat}$ is the length at which on average 50% of individuals are
sexual maturity, and $\delta_{mat}$ is the units of length beyond
$l_{mat}$ at which on average 95% of fish are sexually mature.

Spawning stock biomass at time *t* is then calculated as

@@

where **d^l^** is the recruitment movement matrix

@@

@@

Sigh. I need to work on this section, but I am *le tired*.

A note here Why no reference points? The problem: Fmsy and MSY are
selectivity dependent, and once we allow for different fleets with
different selectivities that critical can move in space, that movement
affects the net selectivity and fishing mortality on the population.

Bmsy is also affected, since if selectivity is very low, you might need
more biomass of immature individuals relative to less biomass of more
mature individuals; a biomass of X can be more or less reproductively
productive depending on the size / age makeup of that biomass.

SSBmsy is more conserved, as over the long term the key things is the
flow of recruits, so we do provide an option for that, but I haven't
messed with it for a while.

There are many different strategies for estimating reference points
conditional on selectivity and regimes, see citations. We leave it to
users to find relevant reference point as required by their methods on
their own, using the sort of MSE examples available in the package.

### MPA Optimization

Our MPA optimization algorithm seems to create an MPA network that
maximizes an objective function, given by

$$
obj_p = \Phi \times scale(\sum_f^F \boldsymbol{\phi_{f,p}}) + (1 - \Phi)  scale(\boldsymbol{\sum_s^S\frac{ssb_s}{ssb0_{s,p}}})
$$ {#eq-obj}

Where *scale* rescales the values to be between 0 and 1 such that the
weight parameter is comparable between vectors. The algorithm works by
first assigning every patch in the system to an MPA. For the first
round, the algorithm then calculates the equilibrium objective function
value resulting from sequentially removing each individual patch from
the MPA network. The cell whose removal from the MPA produces the
highest objective function value is then removed from the MPA network,
and the marginal value of the objective function each of the remaining
cells from the last steps are then stored. For the next steps, we
sampled 25% of the remaining MPA cells with probability proportional to
their marginal contribution to the objective function in the prior
iteration, selecting the cell whose removal produces the highest total
objective function as the cell to remove from the network, and updating
the marginal value of the sampled-but-not-selected cells. This in
essence is a sampling-importance-resampling algorithm. The algorithm
continues until no more cells are left in the MPA. This then provides
for a given objective weighting value $\Phi$ the near-optimal MPA
network at every MPA size from 0 to 100%. We say "near" as they is some
path dependency here, in that for example the network that maximizes the
objective function with 40% of the MPAs does not necessarily contain the
optimal network at 30%. However, the reversed nature of the algorithm,
starting at 100% MPA coverage and working better, produces a more stable
network with regards to this problem than starting at 0% and growing to
100%.

@@

@@

In our coral reef example, we model the dynamics of four species: a
grouper (XX), a shallow-reef snapper (xx), a deep-water snapper (XX),
and a reef shark (XX). Groupers undergo a mass migration to a spawning
aggregation once per year, followed by the sharks. Shallow-reef snappers
stay in reefs closer to shower year-round, while deep-water snappers
stay in the deeper reefs (@fig-coral).

@@

The second fleet primarily targets the deep-water snapper population.
While plate-sized snapper are still prized from this population, we
model fleet two's selectivity as logistic due to high levels of discard
mortality for deep-water snapper resulting from barotrauma (xx). This
fleet catches some grouper, though less than fleet one, and receives no
price for sharks due to the requirements of a certification program
through which they sell their deep-water snapper, though they do
accidentally catch and kill them on occasion. This fleet operates under
open-access dynamics, as fishing is not the primary means of subsistence
and so short-term effort expands and contracts in respond to
profitability of the primarily grouper-driven fishery xx.

```{r}
#| label: fig-coral
#| fig-cap: "Distribution of animals each season under unfished conditions. Numbers show the approximate location of each fleet's port."

@@

patch_biomass %>%
  group_by(critter, step) %>%
  mutate(sbiomass = biomass / max(biomass)) %>%
  ungroup() %>%
  filter(between(step,1,2)) %>% 
  mutate(Season = step) %>% 
  ggplot()+
  geom_tile(aes(x,y, fill = sbiomass),show.legend = FALSE) +
  geom_text_repel(data = coral_ports, aes(x, y, label = fleet), color = "red") +
  facet_grid(Season~critter) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_viridis_c() + 
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank())

@@

We then simulate the impacts of MPAs...

@@

We modeled our pelagic use case loosely on the characteristics of the
Western and Central Pacific Ocean tuna fisheries. We simulate
trajectories of XX species commonly caughts in the region, including
both the highly abundance skipjack tuna (*Katsuwonus pelamis*) to the
heavily depleted oceanic whitetip shark (*Carcharhinus longimanus*). We
pulled publicly available data on catch-per-unit-effort of each of these
species from the WCPO as a proxy for baseline habitat distributions
(Fig.XX). However, for the pelagic case study we simulate a scenario
where climate change is inducing a northerly migration of XX degress per
year in XX species.

These species are targeted by a longline fleet that targets large adult
tunas such as bigeye and yellowfin primarily for high-grade consumption,
and a purse-seine fishery that primarily targets skipjack tunas for bulk
canning. XX. Both fleets are subject to a quota XX add this back in XX.

```{r}
#| label: fig-blue
#| fig-cap: "Distribution of animals under unfished conditions for blue-water simulation."

blue_fauna <- blue_water_fauna_and_fleets$fauna

@@

blue_sim <- simmar(
  fauna = blue_fauna,
  fleets = blue_fleets,
  manager = list(mpas = list(
    locations = mpa_locations,
    mpa_year = 1
  )),
  years = 2
)

patch_biomass <-
  map_df(blue_sim, ~ map_df(.x, ~tibble(biomass = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step") %>%
  mutate(step = as.numeric(step)) %>%
  left_join(grid, by = "patch")

patch_biomass %>%
  group_by(critter, step) %>%
  mutate(sbiomass = biomass / max(biomass)) %>%
  ungroup() %>%
  filter(step == max(step)) %>% 
  ggplot()+
  geom_tile(aes(x,y, fill = sbiomass),show.legend = FALSE) +
  facet_wrap(~critter) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_viridis_c() + 
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank())

@@

#### Key Drivers

The MPA literature points to the ability of a wide range of drivers to
affect the performance of MPAs with regards to conservation and economic
outcomes such as adult and larval dispersal [@gerber2005], the timing of
density dependence [@babcock2011], correlations in space and time across
species [@pons2022; @abbott2012] , fleet dynamics and economic
incentives [@smith2003a; @cabral2019], relationships between size and
fecundity (e.g. hyperallometry) [@marshall2021], heterogeneity in stock
status [@hastings2017; @abbott2012], and many other traits [@gaines2010;
@hilborn2004a].

These studies have all helped guide the science of marine protected
areas, but often focus on modeling the impact of subset of variables of
interest at a time. We demonstrate here how `marlin` can be used to
conduct Monte Carlo simulations across a wide range of drivers to assess
the relative contributions each to the outcomes of protected areas under
a particular set of circumstances.

For this example we model the dynamics of a skipjack tuna *Katsuwonus
pelamis*, a striped marlin *Kajikia audax*, and a oceanic whitetip shark
*Carcharhinus longimanus*. Life history parameters for each species were
pulled from FishLife (@thorson2020), with the exception of steepness,
set to XX, and fecundity. For the tuna and marlin species, fecundity is
assumed to scale with weight modified by the hyperallometric exponent.
For the shark species, we assume a constant value of six pups per year
at all ages greater than the age at maturity. We assume that tunas have
an unfished biomass of XX, marlins XX, and sharks XX, to reflect a
large, medium, and small population.

For the generalized experiments we only include one fishing fleet. We
assign a price per KG of \$XX for tuna, intended to reflect a low value
high volume fishery, \$XX for marlin for a higher value low volume
fishery, and \$0 for sharks to reflect a true bycatch species. We assign
a selectivity of 60% of length at maturity for skipjack to allow for
some overfishing when F/M is high enough, 70% for marlin, and 90% for
shark xx.

We ran `r nrow(state_experiments)` state experiments, varying the state
of nature experienced by a simulation. The candidate states include

| State                  | Distribution          | Description                                     |
|-------------------|-------------------|-----------------------------------|
| $\sigma_{centroid}$    |                       | Standard deviation of habitat centroid          |
| $\sigma_{habitat}$     |                       | Standard deviation around centroid              |
| Spatial catchability   |                       | Catchability varies in space                    |
| Spatial allocation     |                       | Fleet allocation strategy                       |
| F/M                    |                       | Fishing mortality relative to natural mortality |
| Adult diffusion rate   |                       |                                                 |
| Recruit diffusion rate |                       |                                                 |
| Hyperallometry         | $U(1,2)$              | Presence of hyperallometric fec                 |
| DD timing              |                       | Timing of density dependence                    |
| Seasonal Movement      | $\subset[TRUE,FALSE]$ | Seasonal Movement Present                       |

We simulate degrees of species overlap using two parameters, $\sigma_c$
and $\sigma_h$. All species start with the assumption that the centroid
of their preferred habitat occurs in the center-most patch of the two
dimensional system simulated by marlin. For our simulations, we then
generated a random deviation from that default centroid for each species
using $\sigma_c$ in the manner

$$centroid_s \sim N(center,\sigma_c)$$ where *s* denotes species, and
*center* is the coordinate of the central cell. In this way, the larger
$\sigma_c$ grows, the more variation there is between the centroid of
the preferred habitat for each species.

We then control the spread of a creatures preferred habitat around that
centroid per a Gaussian dispersal kernel in the manner

$$habitat_{p,s} = \frac{1}{\sqrt{2\pi\sigma_h^2}}e^{-\frac{d_p^2}{2\sigma_h^2}}$$

Where $d$ is the euclidean distance from the $centroid_s$ to patch $p$
and $\sigma_h$ controls the dispersal of the habitat around that range.

For each of these `r nrow(state_experiments)` state experiments, we then
apply `r nrow(placement_experiments)` placement experiments defined by
an MPA size, an MPA placement strategy, and the number of animals to be
considered in the MPA planning, ranging from one to three, where one
indicates that only one animal is considered in the MPA design process,
while three means that all three species are considered.

>>>>

---
title: "Policy Implications of Coupled Economic and Ecological Dynamics in Aquatic Systems"
author:
    - Daniel Ovando
    - Darcy Bradley
    - Echelle Burns
    - Lennon Thomas
    - James Thorson
date: "`r Sys.Date()`"
format:
  pdf:
    toc: true
    fig-pos: 'H'
  html: default
  docx: default
execute:
  echo: false
  message: false
  warning: false
bibliography: references.bib
params:
  run_name: ["v0.1"]
abstract: "Very rough cut. Communities around the world are increasingly looking to spatial strategies for managing marine ecosystems. However, the design of spatial management strategies faces numerous challenges. Marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond to changes in marine ecosystems in highly complex ways. We present a framework for efficient simulation of the bio-economic dynamics of marine systems, designed to support both management and research. We demonstrate the model's capabilities by focusing on two case studies on the conservation and food security impacts of marine protected areas (MPAs). We highlight how coupled bio-economic dynamics can result in complex trade-offs between conservation, fisheries yields, distributional outcomes in simulated coral reef and plegaic ecosystems. "
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
---

XX explain hyperallometry option

xx add in weird option 5 density dependence

@@

theme_set(theme_classic(base_size = 12, base_family = "Helvetica") + theme(strip.background = element_blank()))

@@

# rdatas <- results_files[str_detect(results_files,'.Rdata')]
# 
# walk(rdatas, ~load(file.path(results_path,.x), envir = .GlobalEnv)) # load results into global environment

@@

`marlin` follows a "models of intermediate complexity" (MICE) philosophy
[@plagányi2014] targeted specifically towards capturing heterogeneous
habitats, movement dynamics, and technical interactions across species.
In its general form, `marlin` simulates the behavior of populations of
spatially explicit biologically independent age-structured animal
populations affected by one or more fishing fleets in time steps
specified by the user (generally annual or seasonal). While `marlin`
does not simulate species interactions, fishing fleets in `marlin` are
capable of targeting and affecting multiple species simultaneously. This
allows the model to simulate processes such as fisheries bycatch or
effort displacement in a way that accounts for the ways that changes in
fleet behavior might affect multiple species in the system. The ability
to track the impacts of fishing fleets across multiple species
simultaneously is particularly important as very few fishing fleets are
truly single species.

@@

A fishing fleet is defined in `marlin` by a set of fishers that can have
different catchability, prices, and selectivity for individual species
within a simulation (each of which we denote as a *metier*). For
example, both a longline and purse seine fleet may capture bigeye
(*Thunnus obesus*) and skipjack (*Katswamus pelamus*) tunas, but the
longline fleet is much more likely to catch bigeye than skipjack, and
*vice versa*. Each fishing fleet (*f*) generates catches or yield (*Y*),
revenues (*R*), costs (*C*) from individual species (*s*) that it
targets in a given time step (*t*) and patch (*p*). The fleet then makes
decisions around how much and where to fish, subject to regulatory
constraints, based on th total profitability across all species in space
and time.

Revenues *R* for fleet *f* in time step *t* are a function of the total
amount caught of each species and its price ($\Pi$). The amount caught
is a function of the selectivity at age of each species for each fleet
($\alpha$), the fishing efficiency of the fleet for that that species
(*q*), the amount of fishing effort of the fleet in question in that
patch in a given time step (*E*), and the total fishing mortality at age
(including all other fleets) for that particular species in that patch
and time (*u*).

@@

*Y* is total catches, which is calculated through the Baranov equation
[@baranov1918], which appropriately accounts for the total instantaneous
mortality (both fishing and natural) and divides the total amount of the
population biomass (*b*) killed between natural (*m*) and fishing (*u*)
sources, the amount of biomass killed through fishing being assigned to yield *Y*.

@@

Selectivity $\alpha$ is modeled through either a logistic form or a dome-shaped form. The logistic form is based on the lengths *l* at which 50% of individuals are selected by the fishing gear, $l^{sel}$, and $\delta$ which is the difference between the length at 50% selectivity and
the length at 95% selectivity.

@@

We approximate the dome-shaped form as a simple normal distribution with mean $l^{sel}$ and standard deviation $\sigma$. The normal density function is re-scaled such that the selectivity is 1 at $l^{sel}$.

$$
\alpha_{a,s,f}= \frac{1}{\sigma_{s,f} \sqrt{2\pi}}e^{-0.5 \left( \frac{l_{a,s} - l^{sel}}{\sigma_{s,f}}  \right)^2}
$$
Lastly  *z* is total mortality, the sum of fishing mortality and natural mortality (*m*).

@@

This means that each species experiences a total mortality at age in a
time step in a patch, individual fractions of which are portioned off as
catches and subsequently revenues for each fleet. Given revenues *R*, we
then calculate profits $\phi$ based on revenues and costs.

@@

Costs *C* are calculated as a function of $\gamma_f$, that the scales the cost per unit effort for fleet *f*,$\beta$ allows for the cost of effort to scale non-linearly, and $\eta$
is an additional cost per unit effort of fishing in each patch *p*,
allowing for the model to account for travel costs for different
patches. Travel costs are calculated based on the euclidean distance of
each patch to the nearest port of a given fleet, users can specify any
number from zero to the number of patches of port locations. When no
ports are specified, travel costs are zero.

@@

@@

The more complex option is "open access". Under open access, the total
amount of effort in a given time step for fleet *f* is a function of the
profitability of that fleet in the proceeding time steps in which
fishing was open, where $\theta$ controls the responsiveness of fleet *f* to the log of
the ratio of total revenues *R* to total costs *C*, and approximates the
percentage amount effort changes in response to a one unit changes in
the log revenue to cost ratio.

@@

When MPAs are implemented, effort that used to operate inside the MPAs can either
leave the fishery, or be redistributed outside the the MPA (the default
behavior). The model can also accept total catch quotas for each species
in each time step. When quotas are activated, if the total catch across
all fleets for a given species under the model's effort dynamics would
exceed the allowable quota for that species, the total effort for each
fleet is decrease proportional to its contribution to the total catch
until the quota is satisfied. As an alternative to catch quotas, users can input an effort cap per fleet, which prevents effort from exceeding a given amount under open-access dynamics, though effort may be reduced below this cap if required by the profit equation. `marlin` also allows users to specify
closed fishing seasons for one or more species in the system. During
closed seasons, no effort is allocated to patches that would result in
capture of the covered species, potentially resulting in zero effort if
the species cannot be avoided. Under these conditions, the amount of
effort in the next open season is a function of the effort and profits
in the last open fishing season.

@@

The underlying population model used is an age structured single-species
model in the manner of @ovando2021a. The population model requires a
large number of parameters. However, if the user supplies either a
scientific or common name for the species in question (scientific
preferred), the model will supply default values for that species based
on the values reported in FishLife [@thorson2020]. All of these values can be replaced with user-supplied values as needed though, the defaults simply ensure that reasonable and internally consistent values for the species in question are supplied as a starting point.

For the population model, numbers *n* at time *t* for age *a* are a function of growth, death, and recruitment

@@

where *BH* is the Beverton-Holt recruitment function, *ssb* is
spawning-stock-biomass. $l_a$ is the mean length at at age, $l_sel$ is the length at which
on average 50% of individuals are selected by the fishery, and
$\delta_{sel}$ are the additional units of length at which on average
95% of fish are selected by the fishery.

Spawning stock biomass *ssb* is calculated by converting age to mean length, calculating weight
at age, maturity at age, and then calculating spawning stock biomass as
the sum of spawning potential at age in a given time step, taking into account the potential for hyperallometry. Age is converted to length through the von Bertalanffy  growth equation, assuming log-normally distributed variation in the length at age.

$$
  l_{a,s} = l_{\infty,s}\left(1 - e^{-k_s(a - a0_{s})}\right)e^{u_s}
$$ {#eq-len}

Where *u* is a log-normally distributed term allowing for variation in
the length at age relationship

$$
u_s \sim normal(0,\sigma_s)
$$ {#eq-sigma}

Biomass *b* at age is then given by the weight at length equation governed by a scaling coefficient $wa_s$ and an exponent $wb_s$ that controls the rate at which volume scales with length

@@

Maturity is then calculated as a logistic function where $l_{mat}$ is the length at which on average 50% of individuals are sexual maturity, and $\delta_{mat}$ is the units of length beyond
$l_{mat}$ at which on average 95% of fish are sexually mature.

$$
  mat_{a,s} = \frac{1}{\left(1 + e^{-log(19)\times\frac{l_{a,s} - lmat_s}{{\delta}mat_s}}\right)}
$$ {#eq-mat)}

Spawning stock biomass at time *t* is then calculated as a function of the numbers at age, the maturity at age, and the weight at age scaled by a parameter $\gamma$. When $\gamma$ is greater than 1, the species is said to experience hyperallometric fecundity, i.e. fecundity increases faster than weight.

@@

where **d^l^** is the recruitment movement matrix (xx make this notation
consistent with the movement block)

@@

5.  Global density dependence allocated by spawning biomass: Density dependence is a function of the
    sum of spawning biomass across all patches, and recruits are then
    distributed according to the distribution of spawning biomass

$$
n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p}}\times{h}\times\sum_{p=1}^P{ssb_{t-1,p}}}{0.2\times{\sum_{p=1}^Pssb0_p}\times(1 - h)+(h - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p}}}\right)\times \frac{ssb_{t-1,p}}{\sum_{p=1}^Pssb_{t-1,p}}
$$ {#eq-dd5}

@@

Fisheries management is often concerned with measuring stock status relative to maximum sustainable yield (MSY) based reference points, though the exact level of stock status desired by societal objectives may vary widely. MS based reference points present a problem for a multi-fleet and spatial-temporal model such as `marlin`. MSY and the fishing mortality rate that would produce MSY, $F_{MSY}$ are a function of fishery selectivity. Fishery selectivity in this model can vary by fleet, and species can be distributed unevenly in both space and time. This means that the net fishing selectivity on a species can vary depending on the dynamics at a given moment, making the definition of an equilibrium concept such as MSY challenging.

As such, we do not report MSY based reference points in the model by default, with the exception of the spawning stock biomass at MSY ($SSB_{MSY}$) which is more conserved across selectivity patterns as it is primarily a function of the mortality and flow of recruits relative to the maturity ogive. There are many different strategies for estimating reference points
conditional on selectivity and regimes, see citations xx. We leave it to
users to find relevant reference point as required by their methods on
their own, using the sort of MSE examples available in the package.

<!-- ### MPA Optimization -->

<!-- Our MPA optimization algorithm seems to create an MPA network that -->
<!-- maximizes an objective function, given by -->

<!-- $$ -->
<!-- obj_p = \Phi \times scale(\sum_f^F \boldsymbol{\phi_{f,p}}) + (1 - \Phi)  scale(\boldsymbol{\sum_s^S\frac{ssb_s}{ssb0_{s,p}}}) -->
<!-- $$ {#eq-obj} -->

<!-- Where *scale* rescales the values to be between 0 and 1 such that the -->
<!-- weight parameter is comparable between vectors. The algorithm works by -->
<!-- first assigning every patch in the system to an MPA. For the first -->
<!-- round, the algorithm then calculates the equilibrium objective function -->
<!-- value resulting from sequentially removing each individual patch from -->
<!-- the MPA network. The cell whose removal from the MPA produces the -->
<!-- highest objective function value is then removed from the MPA network, -->
<!-- and the marginal value of the objective function each of the remaining -->
<!-- cells from the last steps are then stored. For the next steps, we -->
<!-- sampled 25% of the remaining MPA cells with probability proportional to -->
<!-- their marginal contribution to the objective function in the prior -->
<!-- iteration, selecting the cell whose removal produces the highest total -->
<!-- objective function as the cell to remove from the network, and updating -->
<!-- the marginal value of the sampled-but-not-selected cells. This in -->
<!-- essence is a sampling-importance-resampling algorithm. The algorithm -->
<!-- continues until no more cells are left in the MPA. This then provides -->
<!-- for a given objective weighting value $\Phi$ the near-optimal MPA -->
<!-- network at every MPA size from 0 to 100%. We say "near" as they is some -->
<!-- path dependency here, in that for example the network that maximizes the -->
<!-- objective function with 40% of the MPAs does not necessarily contain the -->
<!-- optimal network at 30%. However, the reversed nature of the algorithm, -->
<!-- starting at 100% MPA coverage and working better, produces a more stable -->
<!-- network with regards to this problem than starting at 0% and growing to -->
<!-- 100%. -->

@@

Each of the use cases contains too many parameters and options to be succinctly presented in the text here, and we are concerned that attempting to duplicate all of the parameters of the model here would simply introduce error. As such, readers should consult the accompanying code (provide clear instructions for which files) to view the precise details of each simulation. xx something like this to clarify why we don't have pages of tables with all the parameters and options). Targeted tactical applications though should be sure to consider and document all decisions made around model parameters.

@@

In our coral reef example, we model the dynamics of four species: a
grouper (*Epinephelus fuscoguttatus*), a shallow-reef snapper (*Lutjanus malabaricus*), a deep-water snapper (*Pristipomoides filamentosus*), and
a reef shark (*Carcharhinus amblyrhynchos*). The simulated groupers
undergo a mass migration to a spawning aggregation once per year,
followed by the sharks. Shallow-reef snappers stay in reefs closer to
shower year-round, while deep-water snappers stay in the deeper reefs
(@fig-coral).

@@

The second fleet primarily targets the deep-water snapper population.
While plate-sized snapper are still prized from this population, we
model fleet two's selectivity as logistic due to high levels of discard
mortality for deep-water snapper resulting from barotrauma (xx). This
fleet catches some grouper, though less than fleet one, and receives no
price for sharks due to the requirements of a certification program
through which they sell their deep-water snapper, though they do
accidentally catch and kill them on occasion. This fleet operates under
open-access dynamics, as fishing is not the primary means of subsistence
and so short-term effort expands and contracts in respond to
profitability of the primarily grouper-driven fishery.

Each fleet has a different home-port location, and incurs an additional
cost for fishing each patch in the system in proportion to the euclidean
distance from each patch to each fleet's home port.

We then simulated the outcomes for both food security and conservation
for each of the species and both of the fleets as a function of both MPA
size and MPA design strategy. For this exercise, MPAs are placed with
perfect information and have no design constraints for continuity (i.e.
the selected MPA design can be as patchwork as dictated by the data).

```{r}
#| label: fig-coral
#| fig-cap: "Distribution of fish spawning biomass (color) each season (row) under unfished conditions. The x-axis represents longitude, y-axis latitude. Numbers show the approximate location of each fleet's port. Rows indicate quarterly seasons that repeat each year."

@@

patch_biomass %>%
  group_by(critter, step) %>%
  mutate(sbiomass = biomass / max(biomass)) %>%
  ungroup() %>%
  filter(between(step,1,2)) %>% 
  mutate(Season = step) %>% 
  ggplot()+
  geom_tile(aes(x,y, fill = sbiomass),show.legend = FALSE) +
  geom_text_repel(data = coral_ports, aes(x, y, label = fleet), color = "red") +
  facet_grid(Season~critter) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_viridis_c() + 
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank())

@@

@@

```{r}
blue_fauna <- blue_water_fauna_and_fleets$fauna

@@

blue_resolution <- sqrt(blue_fauna[[1]]$patches)

@@

We modeled our pelagic use case loosely on the characteristics of the
Western and Central Pacific Ocean tuna fisheries. We simulate
trajectories of `r n_distinct(names(blue_fauna))` species commonly
caught in the region, including both the highly abundance skipjack tuna
(*Katsuwonus pelamis*) to the heavily depleted oceanic whitetip shark
(*Carcharhinus longimanus*) (@fig-blue-ssb). We pulled publicly
available data on catch-per-unit-effort of each of these species from
the WCPO as a proxy for baseline habitat distributions (@fig-blue).

These pelagic species are caught by a longline fleet that primarily
targets large adult tunas such as bigeye and yellowfin for high-grade
consumption, and a purse-seine fishery that primarily targets skipjack
tunas for bulk canning. Selectivities were modeled as logistic for the
longline fleet, and dome-shaped for the purse-seine fleet, with
selectivity values pulled from the literature where available (see SI
xx). Both fleets operate under open-access dynamics with an effort cap.
The effort cap was set at the level of effort that resulted in the
desired levels of depletion for each species under open-access dynamics
(@fig-blue), intended to simulate a scenario where managers step in to
prevent further expansion of fishing effort in a fully developed
fishery. For forward-simulation, open-access dynamics can result in
effort decreasing in response to profitability, but cannot result in
effort beyond the effort cap set for each fleet.

For this exercise, we focused on using `marlin` to assess resilience of
the selected "rule of thumb" MPA design strategies to a climate-driven
range shift. Specifically, we simulate an extreme example where the
centroid of each population shifts northward at a rate of 0.2 patches
per year, over a `r experiment_years` time horizon. We designed MPA
networks given the conditions in the starting year, and held that
network constant over the years of the experiment, running one
simulation with and another without the climate-drive range shift. We
then compared effect of this range shift on food security and
conservation outcomes from MPA networks designed based on the pre-range
shift world.

```{r}
#| label: fig-blue-ssb
#| fig-cap: "Trajectory of depletion (spawning biomass divided by unfished spawning biomass) over time prior to implementation of MPAs."

grid <- expand_grid(x = 1:blue_resolution, y= 1:blue_resolution) %>%
  mutate(patch = 1:nrow(.))

mpa_locations <- expand_grid(x = 1:blue_resolution, y = 1:blue_resolution) %>%
  mutate(mpa = TRUE)

blue_sim <- simmar(
  fauna = blue_fauna,
  fleets = blue_fleets,
  years = 50
)

critter_biomass <-
  map_df(blue_sim, ~ map_df(.x, ~tibble(ssb = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step") %>%
  mutate(step = as.numeric(step)) %>%
  left_join(grid, by = "patch") %>% 
  group_by(critter, step) %>% 
  summarise(ssb = sum(ssb)) %>% 
  group_by(critter) %>% 
  mutate(depletion = ssb / ssb[step == 0])

critter_biomass %>% 
  ggplot(aes(step, depletion)) +
  geom_line() + 
  facet_wrap(~critter) +
  scale_y_continuous(limits = c(0, 1))

```{r}
#| label: fig-blue
#| fig-cap: "Distribution of fish spawning biomass under status quo (blue) and range-shifted (red) conditions for blue-water simulation. The x-axis represents longitude, the y-axis latitude. "

blue_climate_sim <- simmar(
  fauna = blue_fauna,
  fleets = blue_fleets,
  habitat = future_habitat,
  years = 20,
  manager = list(mpas = list(
    locations = mpa_locations,
    mpa_year = 1
  ))
)

patch_biomass <-
  map_df(blue_climate_sim, ~ map_df(.x, ~tibble(biomass = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step") %>%
  mutate(step = as.numeric(step)) %>%
  left_join(grid, by = "patch")
